What's in a Domain? Anaylsis of URL Features

### 2023-03-15 paper/paper.tex
Paper update

\begin{abstract}
Many data science problems require processing log data derived from web pages, apis or other
internet traffic sources. URLs are one of the few ubiquitous data fields that describe
internet activity, hence they require effective processing for a wide variety of machine 
learning applications. While URLs are structurally rich, they are not subject
to universal principles of construction making fetaure engineering for internet data an ongoing challenge.

In this research we outline the key categories of information structure in domains and URLs.
We share an open source implementation of these ideas and demonstrate their utility across a range of 
internet traffic machine learning problems. Python package available at https://pypi.org/project/url2features
\end{abstract}

@@

The Uniform Resource Locator (URL) is a ubiquitous element the digital world. 
We use them to advertise and locate businesses, to retrieve news and media and 
to interact with other people across a broad of social applications.
Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed with
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Garera2007,Basnet2012,Basnet2014,Mamun2016,Verma2017,Vazhayil2018,Tupsamudre2019,Li2020}, 
and identification of pages that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising,
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2004,Shih2004,Baykan2009,Meshkizadeh2010,Hernandez2012,Arya2016}, relevance \cite{Kan2005} 
or language\cite{Baykan2013} of webpages. 
Content categorisation has also been applied to spam web pages using URLs alone for the sake
of search results filtering\cite{Chung2010}.

The URL is made up of multiple elements, but at its core is the domain. The domain
contains internal information about both its intended purpose, legitimacy and likely
origin. It is also a source of additional information through requests to Domain Name
Servers (DNS) to understand both its history and place within the network topology.

@@

Common feature engineering strategies exploit these potential sources of information
in ways that are appropriate for a given task. However, while there are wide variety
of approaches that inconsistently applied, there has not yet been a study that evaluates
the overall effectiveness of different strategies across a range of applications.

In this work we develop a feature engineering library that maintains an ontology of URL features 
that relate to underlying structure of the URL, with a method for visualising the importance of features
across the URL structure. These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and efficient feature selection \cite{Basnet2012}. 
We then present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features.

A large amount of recent work on machine learning based on URLs has focused on deep learning approaches that
learn an embedding representation of the content of the URL. Once of the explicit reasons 
for this approach is the potential for sensitivity to the sequential structure of the URL.
\cite{Le2018}

Deep learning approaches include the use of convolutional layers to learn word and character 
embedings \cite{Le2018},

Classify Web pages based on URLs from the 4 Universities WebKB dataset. Comparing features derived from segmentation
of the URL to features derived from either link anchor text or the web page document itself \cite{Kan2004}.

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences and lexical properties \cite{Kan2004,Garera2007,Mamun2016,Tupsamudre2019}, 
creation of lookup tables or likelihood scores based on key sequences
\cite{Meshkizadeh2010}, n-gram or bag of words models \cite{Baykan2009,Verma2017}, 
the generation of task specific embedding vectors\cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information about the domain registration 
and server configuration\cite{Canali2011,Li2020}. In many modern approaches to malicious URL detection
multiple feature engineering approaches are typically combined\cite{Sayamber2014,Li2020}. Increasingly,
sophisticated methods are used to combine, select or learn from combinations of features to adapt to
changing requirements, particularly in internet security applications\cite{Sountharrajan2020,Li2020}.

@@

@@

We design a feature extraction library for URLs that is sensitive to the various sections of URL
structure outlined above. Previous work has presented datasets for specific problems where the features
are provided with a similar structural grouping\cite{Vrban2020}. However, we provide our feature extraction
application designed so that any specific subset of these features can be extracted for a specific task.

@@

The internet security problems appear to derive less utility from the global features when compared to the three
topic classification problems, as is indicated by the strength of the grey lines through the middle of the plot 
for each task. The contributions of the domain name itself and the top level domain vary across the tasks and do
not appear consistant across the top broad categories of secutity and topic classification. Suggesting that developers
need to test these features on a task by task basis rather than relying on heuristics.

@@

We have described the structural elements of a URL from which context specific features can be extracted.
These feature ideas tend to be used indepednetly within different domains. In this paper we have have 
collated them into an open source URL feature generation package and then compare their utility across a range
of indepenent tasks. We have observed that the various structural regions play different roles across 
these tasks, exhibiting varying levels of importance.

>>>>

\begin{abstract}
Many data science problems require processing log data derived from web pages, apis or other
internet traffic sources. URLs are one of the few ubiquitous data fields that describe
internet activity, hence they require effective processing for a wide variety of machine 
learning applications. While URLs are structurally rich, the structure can be both domain specific
and subject to change over time, making feature engineering for URLs an ongoing challenge.

In this research we outline the key structural components of URLs and discuss the information
available within each. We describe methods for generating features on these URL components and 
share an open source implementation of these ideas. In addition, we describe a method for 
exploring URL feature impact that allows for comparison and analysis of the information available inside URLS.
We experiment with a collection of URL classification problems and demonstrate their utility 
of these tools. 
\end{abstract}
% https://pypi.org/project/url2features

@@

The Uniform Resource Locator (URL) is a ubiquitous element the digital world. 
We use them to retrieve news and media, advertise or find businesses and services, 
and to interact with other people across a broad of social applications. Below the
surface many online services use URLs internally for communicating with other digital
services, making the URL a fundamental data point for both users and machines.

Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Garera2007,Basnet2012,Basnet2014,Mamun2016,Verma2017,Vazhayil2018,Tupsamudre2019,Sirigineedi2020,Li2020}, 
and identification of pages that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising,
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2004,Shih2004,Baykan2009,Meshkizadeh2010,Hernandez2012,Arya2016}, relevance \cite{Kan2005} 
or language\cite{Baykan2013} of webpages. 
Content categorisation has also been applied to spam web pages using URLs alone for the sake
of search results filtering\cite{Chung2010,Hassan2008}.

The URL is made up of multiple elements, but at its core is the domain. The domain
contains internal information about both the intended purpose of a URL, its legitimacy 
and the likely country of origin. The domain is also a source of additional information 
through requests to Domain Name Servers (DNS) to understand both its history and structure
 of its network topology and resources.

Beyond the domain, the URL consists of a sequence of words, categories, identfiers, dates
or domain specific abreviations. This sequence can indicate the psychoogical elements of how
information is categeoried, or the internal logic of an application that generates or presents
content dynamically.

@@

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences and lexical properties \cite{Kan2004,Garera2007,Mamun2016,Tupsamudre2019}, 
creation of lookup tables or likelihood scores based on key sequences
\cite{Meshkizadeh2010}, n-gram or bag of words models \cite{Baykan2009,Verma2017}, 
the generation of task specific word embedding vectors based on a segmentation of the URL \cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information about the domain registration 
and server configuration\cite{Canali2011,Li2020}. In many modern approaches to malicious URL detection
multiple feature engineering approaches are typically combined\cite{Sayamber2014,Li2020}. Increasingly,
sophisticated methods are used to combine, select or learn from combinations of features to adapt to
changing requirements, particularly in internet security applications\cite{Sountharrajan2020,Li2020}.

@@

Typically, researchers employ feature engineering strategies exploit these potential sources of information
in ways that are appropriate for a given task. However, while there are wide variety
of approaches that are inconsistently applied, there have been limited studies into
the overall effectiveness of different strategies across a range of applications.

In this work we develop a feature engineering library that maintains an ontology of URL features
related to underlying structure of the URL. In addition we provide a method for visualising the importance of features
across the URL structure. These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and efficient feature selection \cite{Basnet2012}.
We then present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features.

@@

The structural components of a URL shown in Figure \ref{fig:url_structure} are tabulated with an overview of both
the information source in those components are potenital issues with feature engineering in Table \ref{tab:components}

\begin{table*}
\caption{URL Components and Feature Information}
\label{tab:components}
\begin{tabular}{|l|l|l|}
\toprule
Component    &Subcomponent      &Type of Feature Information       \\
\midrule
Protocol     &                  &Type of content, technology, security of transmission             \\
\midrule
Domain       &                  &Textual data indicating general purpose      \\
             &Subdomain         &Specific purpose or categorisation of content           \\
             &Top Level Domain  &Business or service purpose, authenticty, geograpahy, age           \\
\midrule
Path         &                  &Content structure, file or operation system, application type or structure         \\
\midrule
File         &                  &Naming of specific content, psychological intent, executable or static types      \\
\midrule
Parameters   &                  &Substructure of content, personalization, security, tracking         \\
             &Keys              &Set of categories used to differentiate substructure        \\
             &Values            &Set of values for substructure, indication of page content        \\
             &Fragment          &Labels for specifc substructure, source of new application customisation        \\
\bottomrule
\end{tabular}
\end{table*}

We design a feature extraction library for URLs that generates features specific to each component of the
URL indepenently and using an naming convention that permits sepearation and grouping for analysis.
Previous work has presented datasets for specific problems where the features
are provided with a similar structural grouping\cite{Vrban2020}. However, we provide our feature extraction
application designed so that any specific subset of these features can be extracted and analysed 
for a specific task.

@@

The internet security problems appear to derive less utility from the global features when compared to the three
topic classification problems, as is indicated by the strength of the grey lines through the middle of the plot 
for each task. This suggests that these problems rely more heavily on the information from specific components
of the URL structure.

The contributions of the domain remain high across all problems evaluated. However, whether the key elements is the
domain name itself, the subdomain or the top level domain varies considerably across the tasks and does
not appear consistant across the top broad categories of security and topic classification problems. 
This suggests that data scientists and machine learning engineers
need to test these features on a task by task basis rather than relying on heuristics.

@@

We have described the structural elements of a URL from which context specific features can be extracted.
These features tend to be used indepedently within different problem types and tuned for specific purposes. 
In this paper we have described an open source URL feature generation package that is both sensitive to the 
source of information within each component and sufficiently flexible for general use. We conducted experiments
using the features across a range of indepenent URL classification tasks.

Our experiments demonstrated that the various structural regions play different roles across 
these tasks, exhibiting varying levels of importance. Some patterns (like the strength of global URL features)
could be attributed to broader categories of the task, while others appear task specific.
We have observed that the domain name, with its subcomponents, is a strong and consistent contributor to the
performance of models across tasks. The source of information from within the domain varies, but it remains
a rich source of information for machine learning tasks that rely on the URL as acentral feature.
